
---
title: "5330_Assignment 2"
output:
  pdf_document: default
  html_document: default
date: "2026-02-19"



CH7
```{r}
# --------------------------------------------------
# Question 1
# Load required packages and dataset
# --------------------------------------------------

library(tidyverse)

# Load the GSS dataset
load("/Users/Lcc/gss2018.rda")

# View object names
ls()

```

```{r}
# --------------------------------------------------
# Question 2
# Select only USETECH and MARITAL variables
# --------------------------------------------------

data <- GSS %>%
  select(USETECH, MARITAL)

# Check structure
str(data)

```

```{r}
# --------------------------------------------------
# Question 3
# Clean USETECH variable
# Convert -1, 998, and 999 to NA
# --------------------------------------------------

data <- data %>%
  mutate(
    USETECH = na_if(USETECH, -1),
    USETECH = na_if(USETECH, 998),
    USETECH = na_if(USETECH, 999)
  )

summary(data$USETECH)

```
Interpretation:
These invalid codes are now treated as missing values.

```{r}
# --------------------------------------------------
# Question 4
# Clean MARITAL variable
# Convert 9 to NA and label categories
# --------------------------------------------------

data <- data %>%
  mutate(
    MARITAL = na_if(MARITAL, 9),
    MARITAL = factor(
      MARITAL,
      levels = c(1, 2, 3, 4, 5),
      labels = c(
        "Married",
        "Widowed",
        "Divorced",
        "Separated",
        "Never married"
      )
    )
  )

str(data$MARITAL)
table(data$MARITAL, useNA = "ifany")

```
Interpretation:
MARITAL is now a categorical factor variable.

```{r}
# --------------------------------------------------
# Question 5
# Exploratory Data Analysis
# Examine distribution of variables
# --------------------------------------------------

summary(data$USETECH)
table(data$MARITAL)

```
Interpretation:
This shows the distribution of technology use and marital categories.

```{r}
# --------------------------------------------------
# Question 6
# Visualize technology use by marital status
# --------------------------------------------------

ggplot(data, aes(x = MARITAL, y = USETECH)) +
  geom_boxplot() +
  labs(
    title = "Technology Use by Marital Status",
    x = "Marital Status",
    y = "Technology Use"
  )

```
Interpretation:
The boxplot allows comparison of medians and spread across groups.

```{r}
# --------------------------------------------------
# Question 7
# Calculate mean and standard deviation
# Drop NA values before calculation
# --------------------------------------------------

descriptives <- data %>%
  drop_na(USETECH, MARITAL) %>%
  group_by(MARITAL) %>%
  summarize(
    mean_use = mean(USETECH),
    sd_use = sd(USETECH)
  )

descriptives

```
Interpretation:
This shows average technology use and variability by marital group.

```{r}
# --------------------------------------------------
# Question 8
# Conduct one-way ANOVA
# Test whether technology use differs by marital status
# --------------------------------------------------

anova_data <- data %>%
  drop_na(USETECH, MARITAL)

anova_model <- aov(USETECH ~ MARITAL, data = anova_data)

summary(anova_model)
```
``Interpretation:
If p < .05, there is a significant difference in technology use across marital groups.`

```{r}
# --------------------------------------------------
# Question 9
# Perform Tukey post-hoc test
# Identify which groups differ
# --------------------------------------------------

TukeyHSD(anova_model)

```
Interpretation:
Significant pairwise comparisons indicate which marital groups differ in technology use.




CH 8

```{r}
# --------------------------------------------------
# Question 1
# Load required packages and Chapter 8 dataset
# --------------------------------------------------

library(tidyverse)

data_ch8 <- read.csv("/Users/Lcc/water_educ_2015_who_unesco_ch8.csv")

str(data_ch8)


```

```{r}
# --------------------------------------------------
# Question 2
# Select variables for analysis:
# male.in.school, female.in.school, perc.basic2015water
# --------------------------------------------------

data_ch8 <- data_ch8 %>%
  select(male.in.school, female.in.school, perc.basic2015water)

summary(data_ch8)

```

```{r}
# --------------------------------------------------
# Question 3
# Examine distribution of male.in.school
# --------------------------------------------------

summary(data_ch8$male.in.school)

ggplot(data_ch8, aes(x = male.in.school)) +
  geom_histogram(bins = 20)

```
Interpretation:
This shows the distribution of male school enrollment across countries. The histogram allows us to examine skewness and overall spread.

```{r}
# --------------------------------------------------
# Question 4
# Examine distribution of female.in.school
# --------------------------------------------------

summary(data_ch8$female.in.school)

ggplot(data_ch8, aes(x = female.in.school)) +
  geom_histogram(bins = 20)

```
Interpretation:
This shows how female school enrollment is distributed across countries and whether the variable appears normally distributed.

```{r}
# --------------------------------------------------
# Question 5
# Examine distribution of perc.basic2015water
# --------------------------------------------------

summary(data_ch8$perc.basic2015water)

ggplot(data_ch8, aes(x = perc.basic2015water)) +
  geom_histogram(bins = 20)

```
Interpretation:
This histogram shows the distribution of access to basic water services across countries.


```{r}
# --------------------------------------------------
# Question 8 (Assumption Check)
# Scatterplot: male.in.school vs female.in.school
# --------------------------------------------------

ggplot(data_ch8, aes(x = male.in.school, y = female.in.school)) +
  geom_point() +
  geom_smooth(method = "lm")

```
Interpretation:
The scatterplot allows us to assess linearity. If the relationship appears roughly linear, regression assumptions may be met.

```{r}
# --------------------------------------------------
# Question 6
# Linear regression: female.in.school ~ male.in.school
# --------------------------------------------------

model1 <- lm(female.in.school ~ male.in.school, data = data_ch8)

summary(model1)

```
Interpretation:
If p < 0.05, male enrollment significantly predicts female enrollment. The R-squared indicates how much variation in female enrollment is explained by male enrollment.

```{r}
# --------------------------------------------------
# Residual diagnostics
# --------------------------------------------------

plot(model1)

```
Interpretation:
Residual plots help assess normality and homoscedasticity. If assumptions are reasonably satisfied, linear regression is appropriate.

```{r}
# --------------------------------------------------
# Question 8 (Assumption Check for second pair)
# Scatterplot: male.in.school vs perc.basic2015water
# --------------------------------------------------

ggplot(data_ch8, aes(x = male.in.school, y = perc.basic2015water)) +
  geom_point() +
  geom_smooth(method = "lm")

```
Interpretation:
This plot helps assess whether access to water is linearly related to male school enrollment.

```{r}
# --------------------------------------------------
# Question 6
# Linear regression: perc.basic2015water ~ male.in.school
# --------------------------------------------------

model2 <- lm(perc.basic2015water ~ male.in.school, data = data_ch8)

summary(model2)

```
Interpretation:
If p < 0.05, male school enrollment significantly predicts access to water. The strength of association is indicated by R-squared.

```{r}
# --------------------------------------------------
# Residual diagnostics
# --------------------------------------------------

plot(model2)

```
Interpretation:
If residuals appear randomly distributed, regression assumptions are reasonably met. If not, a nonparametric approach (Spearman correlation) may be more appropriate.

```{r}
# --------------------------------------------------
# Question 9
# Nonparametric alternative (Spearman correlation)
# --------------------------------------------------

cor.test(data_ch8$male.in.school,
         data_ch8$female.in.school,
         method = "spearman")

cor.test(data_ch8$male.in.school,
         data_ch8$perc.basic2015water,
         method = "spearman")

```
Interpretation:
If assumptions are violated, Spearman correlation evaluates monotonic relationships without requiring normality.
